Layered Formal Verification of an Industrial TCP/IP Stack

### 2021-06-02 SecDev21/paper.tex
adding KLEE citation to save space

The original \emph{CycloneTCP} library provides no guarantees that the implementation respects the TCP functional specification. Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit. We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}. The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions. To achieve this, the rest of the TCP protocol, not related to the user task, also has to be taken into account. The reason for this is two-fold. First, although user-functions can trigger state transitions, the actual transitions are done by library functions that do not belong to the user-task set of functions. Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user-related function can exhibit. The TCP user API-related functions were fully translated to SPARK. SPARK bindings were also provided to most of the rest of the library's C functions to allow their invocation from the SPARK code.

@@

Symbolic execution is a means of executing programs with symbolic values rather than concrete values \cite{symbolic_exec_survey:2018}. Exhaustive symbolic execution can be considered sound and complete (it prevents false negatives and false positives), and thus, it is complementary to formal verification. In \cite{vanoverberghe2008using}, the authors present how symbolic execution can be used to improve deductive verification, and in \cite{kassios2012comparing} the authors demonstrated the technique's value in inferring contracts. For our work, the KLEE\footnote{\url{https://klee.github.io}} symbolic execution engine is used. KLEE is built upon the LLVM optimizer's bytecode, and it offers a very simple yet powerful C interface.

>>>>

The original \emph{CycloneTCP} library provides no guarantees that the implementation respects the TCP functional specification. Thus, there is no assurance that the implementation will allow only valid transitions between the different states that a TCP session can exhibit. We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}. The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions. To achieve this, the rest of the TCP protocol, not related to the user task, also has to be taken into account. The reason for this is two-fold. First, although user-functions can trigger state transitions, the actual transitions are done by library functions that do not belong to the user-task set of functions. Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user-related function can exhibit. The TCP user API-related functions were fully translated to SPARK. SPARK bindings were also provided to most of the rest of the library's C functions to allow their invocation from the SPARK code.

@@

Symbolic execution is a means of executing programs with symbolic values rather than concrete values \cite{symbolic_exec_survey:2018}. Exhaustive symbolic execution can be considered sound and complete (it prevents false negatives and false positives), and thus, it is complementary to formal verification. In \cite{vanoverberghe2008using}, the authors present how symbolic execution can be used to improve deductive verification, and in \cite{kassios2012comparing} the authors demonstrated the technique's value in inferring contracts. For our work, the KLEE v2.1 symbolic execution engine is used \cite{KLEE}. KLEE is built upon the LLVM optimizer's bytecode, and it offers a very simple yet powerful C interface.
